### 2016-07-13 capability-machine.tex
Some remarks on symbolic malloc spec.
Another remark.

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it follows the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc v.2]
  \begin{align*}
    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments\ldotp\\
    &\quad \forall n, i, \size \in \nats\ldotp\forall b,e,a,m_b,m_e\in \Addrs\ldotp\forall p \in \Perms \ldotp\\
    &\qquad W_{\malloc} = [i \mapsto \iota_{\malloc}] \land \\
    &\qquad \dom(\hsfoot) = [m_b,m_e] \land \\
    &\qquad \memheap = \hsfoot \uplus \hsframe \land\\
    &\qquad \heapSat[\hsfoot]{n}{W_{\malloc}} \land \\
    &\qquad \memreg(r_1) = \size \land \\
    &\qquad \size \geq 0 \land \\
    &\qquad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
    &\qquad\Rightarrow \\
    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot', \hsframe' \in \HeapSegments\ldotp\\
    &\qquad \qquad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \\
    &\qquad \qquad \quad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe' \land\\
    &\qquad \qquad \quad W' = [i \mapsto \iota_{\malloc}, k \mapsto \iota^0_{b',e'}] \land \\
    &\qquad \qquad \quad \heapSat[\hsfoot']{n-j}{W'} \land \\
    &\qquad \qquad \quad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \quad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \quad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
\end{align*}
\end{specification}
\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}

>>>>

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it follows the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc v.2]
  \begin{align*}
    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments\ldotp\\
    &\quad \forall n, i, \size \in \nats\ldotp\forall b,e,a,m_b,m_e\in \Addrs\ldotp\forall p \in \Perms \ldotp\\
    &\qquad W_{\malloc} = [i \mapsto \iota_{\malloc}] \land \\
    &\qquad \dom(\hsfoot) = [m_b,m_e] \land \\
    &\qquad \memheap = \hsfoot \uplus \hsframe \land\\
    &\qquad \heapSat[\hsfoot]{n}{W_{\malloc}} \land \\
    &\qquad \memreg(r_1) = \size \land \\
    &\qquad \size \geq 0 \land \\
    &\qquad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
    &\qquad\Rightarrow \\
    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot', \hsframe' \in \HeapSegments\ldotp\\
    &\qquad \qquad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \\
    &\qquad \qquad \quad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe' \land\\
    &\qquad \qquad \quad W' = [i \mapsto \iota_{\malloc}, k \mapsto \iota^0_{b',e'}] \land \\
    &\qquad \qquad \quad \heapSat[\hsfoot']{n-j}{W'} \land \\
    &\qquad \qquad \quad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \quad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \quad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
\end{align*}
\end{specification}
\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
  pc unspecified as it is not important for malloc users?}
\dominique{Resulting frame should be equal to original.}
\dominique{Allocated pointer should have the right size.}
\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}
